Demystifying the Computer Science \\ PhD Admission in the US\\{\large A Handbook for International \emph{and Domestic} Students

### 2024-03-17 demystify.tex
updates

Another effective way is taking graduate level courses in the topics you are interested in.  Professors teaching these \emph{special topics courses} and \emph{research seminars} might be recruiting students---giving you even a higher chance. Do well in the class, answer questions, talk to the prof after classes, etc---being stand out.  Many professors, including myself, prefer taking in new students this way.  It gives both the professor and student more time, e.g., a whole semester, to work and evaluate the relationship before making any commitment (sounds a bit like a marriage!).
You can also ask if you can do an independent study or research with a prof. This can be informal (no credit) and takes place during the summer or winter break.  For example, I do this with multiple students, many of whom are undergrads. Many will drop because they find they don't like my research, but some will stay and become my PhD students.

Ultimately, choose a prof. that fits you the most by communicating with them, taking their courses, meeting and asking them questions, and talking to their current students. It will take time and effort, but since you will be working with this person for 5+ years, it is important to try to find the right one.

@@

In many cases you do not need to do a PhD in CS to work in your area of interest. For example, in addition to a traditional CS department, GMU has IST and Cybersecurity departments, both of which have faculty  with PhD in CS and work on CS topics (e.g., AI, Security, Robotics).  So it is totally possible that you still get to do CS research and publish in CS-focused venues even if you're not in a traditional CS program.  It is  common to see faculty with CS PhD in a non-CS department as well as faculty with non-CS PhD in CS department.

>>>>

Another effective way is taking graduate level courses in the topics you are interested in.  Professors teaching these \emph{special topics courses} and \emph{research seminars} might be recruiting students---giving you even a higher chance. Do well in the class, answer questions, talk to the prof after classes, etc---being stand out.  Many professors, including myself, prefer taking in new students this way.  It gives both the professor and student more time, e.g., a whole semester, to work and evaluate the relationship before making any commitment (sounds a bit like a marriage!).

You can also ask if you can do an independent study or research with a prof. This can be informal (no credit) and takes place during the summer or winter break.  For example, I do this with multiple students, many of whom are undergrads. Many will drop because they find they don't like my research, but some will stay and become my PhD students.

Ultimately, choose a prof. that fits you by communicating with them, taking their courses, meeting and asking them questions, and talking to their current students. It will take time and effort, but since you will be working with this person for 5+ years, it is important to try to find the right one.

@@

You \emph{do not} need to do a PhD in CS to work in your area of interest. For example, in addition to a traditional CS department, GMU has IST and Cybersecurity departments, both of which have faculty  with PhD in CS and work on CS topics (e.g., AI, Security, Robotics).  So it is totally possible that you still get to do CS research and publish in CS-focused venues even if you're not in a traditional CS program.  It is  common to see faculty with CS PhD in a non-CS department as well as faculty with non-CS PhD in CS department.
